---
title: "Lab 4"
author: "Damodar Pai"
date: "2024-02-08"
output: html_document 
  toc: yes 
  toc_float:yes
---
#Tasks 

##Task 1
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
getwd() 
```

## Task 2

```{r}
spruce = read.csv("SPRUCE.csv") 
tail(spruce)
```
 
## Task 3 
```{r} 
library(s20x) 
trendscatter(Height ~ BHDiameter, f=0.5,spruce) 

``` 
The above plot is the fitting of a line to the spruce data with f=0.5 level smoothness. 

```{r}
spruce.lm = with(spruce, lm(Height ~ BHDiameter))  
height.res = residuals(spruce.lm)  
height.fit = fitted(spruce.lm)  
```
Above we are creating the spruce model object and then using that we are creating the residual and fitted objects. 
```{r}
plot(height.fit, height.res, col="R") 
``` 

```{r}
trendscatter(height.fit, height.res) 
```
The plot is in a uni modal and symmetric shape. The above data is the residuals vs the fitted data. This basically tells us how close our model is to the given data from a linear moidel. We are trying to create a model so that the residual heights are as close to 0 as possible.  
This curve is very similar to the previous trendscatter plot. 
```{r}
plot(spruce.lm) 
```
```{r}
normcheck(spruce.lm,shapiro.wilk = TRUE) 

```
There clearly isn't constant variance as we can tell from our previous trendscatter plot. Thus, it seems like it doesn't make sense to apply a straight line and we should explore other types of models including a quadratic model. 

```{r}
quad.lm = 
```




## Task 4 

