---
title: "Lab1"
author: "Damodar Pai"
date: "r Sys.Date()"
output:
  html_document:
    toc: yes 
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks 

Lab 1 tasks 

## Task 1 
```{r}
getwd()
```
## Task 2 
```{r}
ddt <- read.csv("DDT.csv") 
head(ddt) 

```
## Task 3 
```{r} 
sapply(ddt,class)   
```
Qualitative variables are "River" and "Species" because their data types are character. 
Quantitative variables are "Mile", "Length", "Weight", and "DDT" are either numeric or integers meaning that they're numbers. 

```{r} 
table(ddt$SPECIES) 
```
There are 3 species. 

```{r}
ddt[ddt$SPECIES == "LMBASS" & ddt$WEIGHT > 800,]  
ddt[ddt$RIVER == "SCM" & ddt$DDT > 4.0,] 
ddt[ddt$RIVER == "SCM" & ddt$DDT > 4.0,] 
with(ddt, ddt[SPECIES == "LMBASS" & WEIGHT > 800,]) 
?which.max()
```
ddt[]

  



## Task 4 
```{r} 
L <-ddt$LENGTH 
mean(L)  
```
Q.1) The mean is 42.8125 cm.
```{r}
W <-ddt$WEIGHT 
sd(W) 
```
Q.2) The standard deviation is 376.5461 gms.  
```{r}
library(dplyr)
library(ggplot2) 
with(ddt,plot(WEIGHT,LENGTH))  
#plot(ddt$WEIGHT,ddt$LENGTH, xlab = "WEIGHT",ylab="LENGTH")
```
Q.3) 0 - No(It seems like the x-axis is Length and the y-axis is Weight in the give graph) 
```{r}
v = 1:20  
v/20
```

Q.4) The last value of v/20 is 1. 

## Task 5
```{r}
riv = table(ddt$RIVER) 
riv
barplot(riv,col= rainbow(4))
rivAndSp = with(ddt,table(RIVER,SPECIES))  
barplot(rivAndSp, col=rainbow(4))

```


## Task 6
```{r}
sp = with(ddt,table(SPECIES))
pie(sp, col=rainbow(3)) 
pie(riv, col=rainbow(4))
```


## Task 7
```{r}
with(ddt,boxplot(DDT,ylab="DDT",col="Blue")) 
with(ddt,boxplot(WEIGHT,ylab="WEIGHT",col="Blue")) 
with(ddt,boxplot(LENGTH,ylab="LENGTH",col="Blue")) 


```



## Task 8
```{r}
rivcol=with(ddt, ifelse(RIVER=="FCM","Red",
                        ifelse(RIVER=="LCM","Blue",
                               ifelse(RIVER=="SCM","Green","Black"))))
fishcol=with(ddt,ifelse(SPECIES=="CCATFISH","Red",
                        ifelse(SPECIES=="SMBUFFALO","Blue","Green")))

coplot(LENGTH~WEIGHT|RIVER,data=ddt,col=rivcol)
coplot(DDT~WEIGHT|SPECIES,data=ddt,col=fishcol) 

```


## Task 9
```{r}
library(ggplot2) 
ggplot(ddt,mapping = aes(SPECIES,WEIGHT,fill=RIVER)) + geom_boxplot() +ggtitle("Damodar Pai") 
ggplot(ddt,mapping = aes(RIVER,LENGTH,fill=SPECIES)) + geom_violin() +ggtitle("Damodar Pai") 
ggplot(ddt,mapping = aes(WEIGHT,LENGTH,color=SPECIES)) + geom_point() +ggtitle("Damodar Pai") 

```


